A thesis draft

### 2018-11-09 res/sections/08-Background.tex
80 columns

\subsection{The SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption{SDN architecture}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted into three different layers, based on their
functionalities. \emph{SDN Applications} are programs that communicate the
behaviour to the SDN controllers and request for needed resources and data
through APIs, often called \emph{northbound interfaces}(NBI). NBIs add a level
of abstraction to the higher level application on the set of functionalities of
the controllers: ideally, taking advantage of NBIs, application could use
controllers of different vendors in the same way. So, SDN applications can be
seen as 2 layers programs, the logic and the communication part that uses NBIs.
Theese applications can also build an abstracted view of the network, using
information of controllers, in order to make easier the analysis of the traffic
and simplifing the decision making process.

@@

\subsection{Network Virtualization}
%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAccessKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41fBNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_network_virtualization.pdf
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hypervisor, decupling the logical
network from the harwdare one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastucture
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service. Most known instances of NV are \emph{VPN},
dedicated networks that allows the communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of netvork virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, tipically
implemented at application layer.

>>>>

\subsection{The SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption{SDN architecture}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted in three different layers, based on their
functionalities. \emph{SDN Applications} are programs that communicate the
behaviour to the SDN controllers and request for needed resources and data
through APIs, often called \emph{northbound interfaces}(NBI). NBIs add a level
of abstraction to the higher level application on the set of functionalities of
the controllers: ideally, taking advantage of NBIs, application could use
controllers of different vendors in the same way. So, SDN applications can be
seen as 2 layers programs, the logic and the communication part that uses NBIs.
Theese applications can also build an abstracted view of the network, using
information of controllers, in order to make easier the analysis of the traffic
and simplifing the decision making process.

@@

\subsection{Network Virtualization}
%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
%r k _virtualization.pdf
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hypervisor, decupling the logical
network from the harwdare one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastucture
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service. Most known instances of NV are \emph{VPN},
dedicated networks that allows the communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of netvork virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, tipically
implemented at application layer.
